---
title: "Flexible Model Parameters"
author: "Sara Michele Schaal"
date: "6/13/2019"
output: html_document
---

```{r}
w.zik <- function(zik, theta, omega.sq){
  1 - (zik - theta)^2/omega.sq
}

```

For the 2-patch model, we assume that each patch has an optimum of +1 and -1.  We assume that for a given number of loci that affect the trait(`ntot`) at the lowest level of redundancy, their effect sizes (`alpha`) are $1/(2*ntot)$.  For this, we can use the supplemental equation from Yeaman 2015 Am Nat to calculate the critical migration rate $m_c$.  At $m>m_c$, alleles are prone to swamping by migration.

In our SLiM simulation, we can simulate up to 2000 SNPs of equal effect sites.

```{r}

theta.1 <- 1
theta.2 <- -1
omega.sq <- 25
ntot <- c(2,10,20,50,100,500)
alpha <- 1/(ntot)

m_crit <- function(alpha, theta.1, theta.2, omega.sq, N){
  w1bb <- 1#w.zik(4*alpha,theta.1, omega.sq)/
    #w.zik(4*alpha, theta.1, omega.sq)
  w1Bb <- w.zik(2*alpha,theta.1, omega.sq)/
    w.zik(4*alpha, theta.1, omega.sq)
  w1BB <- w.zik(0, theta.1,omega.sq)/
    w.zik(4*alpha, theta.1, omega.sq)
  
  w2bb <- w.zik(4*alpha,theta.2, omega.sq)/
    w.zik(0, theta.2, omega.sq)
  w2Bb <- w.zik(2*alpha,theta.2, omega.sq)/
    w.zik(0, theta.2, omega.sq)
  w2BB <- 1#w.zik(0, theta.2,omega.sq)/
    #w.zik(0, theta.2, omega.sq)
  
  w1bb;w1Bb; w1BB
  w2bb;w2Bb; w2BB
  return(1/(w1Bb/(w1Bb-w1BB*(1+(1/(4*N))))-w2Bb/(w2BB*(1+(1/(4*N)))-w2Bb))
  )
}

N=1000
mc <- m_crit(alpha, theta.1, theta.2, omega.sq, N)
cbind(ntot, alpha, mc=round(mc,5), lowmc=0.0100*round(mc,5), highmc=100*round(mc,5))

```