---
title: "processInversionSims"
author: "Sara Michele Schaal"
date: "6/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load Packages and Read in parameters ###
```{r}
library(ggplot2)
library(dplyr)
library(purrr)
options(scipen = 999)

# Load File Path
folder <- "results/Inversion/20201115_FullSet"

# upload a file with all the simulations that were run in this folder
df.params <- read.table("src/InvSimParams.txt", header = TRUE)
```

### Run the following code chunk once to get full data files to do further analyses on ###
```{r}
## Inversion Through Time
df.invTime <- NULL
count <- 0
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  invTimeNewFile <- read.table(paste(folder, seed, "_outputInvTime.txt", sep=""), header = TRUE, stringsAsFactors = FALSE)
  if(nrow(invTimeNewFile) > 0){
    invTimeNewFile$seed <- seed
    df.invTime <-  rbind(df.invTime, invTimeNewFile)
  }
  count <- count + 1
  print(count)
}
write.table(df.invTime, "FullSet_invTime.txt", row.names = FALSE)

## Inversion Summary Data
df.invData <- NULL
count <- 0
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  invData <- read.table(paste(folder, seed, "_outputInvSumInfo.txt", sep=""), header = TRUE,
                        stringsAsFactors = FALSE)
  if(nrow(invData) > 0){
    invData$seed <- seed
    df.invData <- rbind(df.invData, invData)
  }
  count <- count + 1
  print(count)
}
write.table(df.invData, "FullSet_invData.txt", row.names = FALSE)

## QTN Inversion Through Time
df.invQTNTime <- NULL
count <- 0
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  invQTNTimeNewFile <- read.table(paste(folder, seed, "_outputInvQtn.txt", sep=""), header = TRUE, stringsAsFactors = FALSE)
  if(nrow(invQTNTimeNewFile) > 0){
    invQTNTimeNewFile$seed <- seed
    df.invQTNTime <-  rbind(df.invQTNTime, invQTNTimeNewFile)
  }
  count <- count + 1
  print(count)
}
write.table(df.invQTNTime, "FullSet_invQTNTime.txt", row.names = FALSE)

## QTN Inversion Summary Data
df.invQTNData <- NULL
count <- 0
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  invQTNDataNewFile <- read.table(paste(folder, seed, "_outputInvQtnSumInfo.txt", sep=""), header = TRUE, stringsAsFactors = FALSE)
  if(nrow(invQTNDataNewFile) > 0){
    invQTNDataNewFile$seed <- seed
    df.invQTNData <-  rbind(df.invQTNData, invQTNDataNewFile)
  }
  count <- count + 1
  print(count)
}
write.table(df.invQTNData, "FullSet_invQTNSumInfo.txt", row.names = FALSE)

## Pop Dynamics
df.popDyn <- NULL
count <- 0
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  popDynNewFile <- read.table(paste(folder, seed, "_outputPopDynam.txt", sep=""), header = TRUE, stringsAsFactors = FALSE)
  if(nrow(popDynNewFile) > 0){
   popDynNewFile$seed <- seed
   df.popDyn <-  rbind(df.popDyn, popDynNewFile)
  }
  count <- count + 1
  print(count)
}
write.table(df.popDyn, "FullSet_popDyn.txt", row.names = FALSE)

## SimStats
df.simStats <- NULL
count <- 0
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  simStatsNewFile <- read.table(paste(folder, seed, "_outputSimStats.txt", sep=""), stringsAsFactors = FALSE)
  simStatsNewFile$seed <- seed
  df.simStats <-  rbind(df.simStats, simStatsNewFile)
  count <- count + 1
  print(count)
}
df.simStats <- df.simStats[,2:ncol(df.simStats)]
colnames(df.simStats) <- c("mig1", "mig2", "pop1N", "pop2N", "mu_base", "mu_inv", "r", "alpha", "sigmaK", "burnin", "dom", "enVar", "Seed")
write.table(df.simStats, "FullSet_simStats.txt", row.names = FALSE)

## Mutations File
df.finalMuts <- NULL
count <- 0
no.Data <- NULL
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  if(file.exists(paste(folder, seed, "_outputMutations.txt", sep=""))){
    finalMutsNewFile <- read.table(paste(folder, seed, "_outputMutations.txt", sep=""), stringsAsFactors = FALSE, header = TRUE)
  finalMutsNewFile$seed <- seed
  df.finalMuts <-  rbind(df.finalMuts, finalMutsNewFile)
  } else {
    no.Data <- c(no.Data, seed)
  }
  count <- count + 1
  print(count)
}

write.table(no.Data,"seedsToRerun.txt", row.names = FALSE)
colnames(df.finalMuts)[2] <- "mut_id" 
file.exists(paste(folder, seed, "_outputMutations.txt", sep=""))
write.table(df.finalMuts, "FullSet_finalMuts.txt", row.names = FALSE)


```

## Unique Parameters
Subset original parameters dataframe to identify all unique parameter combinations. Then using those parameter values identify the simulation seeds of replicate simulations for each set of parameters.
```{r}

unique.params <- unique(df.simStats[c("mig1","mig2", "pop1N", "pop2N", "mu_base", "mu_inv", "r", "alpha", "sigmaK", "burnin", "dom", "enVar")])
reps <- 2
for(i in 1:reps){
  seedColName <- paste("Seed", i, sep = "")
  unique.params[,seedColName] <- NA
}

for(i in 1:nrow(unique.params)){
  for(j in 1:nrow(df.simStats)){
    if(apply(unique.params[i,1:12], 1, paste, collapse = " ") == 
       apply(df.simStats[j,c("mig1","mig2", "pop1N", "pop2N", "mu_base", "mu_inv", "r", "alpha", 
                             "sigmaK", "burnin", "dom", "enVar")], 1, paste, collapse = " ")){
      for(k in 1:reps){
        seedCol <- paste("Seed", k, sep="")
        if(is.na(unique.params[i, seedCol])){
          unique.params[i, seedCol] <- df.simStats$Seed[j]
          break
        } 
      }
    }
  }
}

```

### Average Replicate Simulations - Pop dynamics file ONLY NEED THIS CHUNK ONCE
Create a dataframe that has the average of all the replicate parameter values. Then bind the corresponding parameters to the dataframe for ease of plotting and analysis later. 
```{r}

########################
### average for loop ###
###################################################################

df.LA.average <- NULL
df.LA.stderr <- NULL
df.LA.all <- NULL
count <- 0
# step through each line of the unique parameters dataframe
for(i in 1:nrow(unique.params)){
    ptm <- proc.time()
    # create empty variable for putting each data set that should be average
    df.average <- NULL
    av.seeds <- NULL
    # step through the population dynamics dataframe
    for(j in 1:nrow(df.popDyn)){
      # step through the different seeds that have the unique parameters
      for(k in 1:reps){
        seedCol <- paste("Seed", k, sep="")
        # if that seed is not an NA then do the next step
        if(!is.na(unique.params[i, seedCol])){
          # if the seed from unique parameters matches the seed of the pop dynam dataset store it
          if(df.popDyn$seed[j] == unique.params[i, seedCol]){
            # bind data from same parameter combination together for averaging later
            df.average <- rbind(df.average, df.popDyn[j,])
            # store unique seeds to this dataset
            av.seeds <- unique(c(av.seeds, unique.params[i, seedCol]))
          }
        }
      }
    }
    
    
    ## average columns of interest
    # df.averageSubset <- df.average[,vect.aggCols]
    new.average <- aggregate(.~sim_gen, data = df.average, FUN = mean)
    SE <- function(x){
      sd(x)/sqrt(length(x))
    }
    new.se <- aggregate(.~sim_gen, data = df.average, FUN = SE)
    df.simParams <- data.frame(mu_inv = rep(unique.params$mu_inv[i], nrow(new.average)), 
                               mig = rep(unique.params$mig1[i], nrow(new.average)),
                               alpha = rep(unique.params$alpha[i], nrow(new.average)), 
                               sigmaK = rep(unique.params$sigmaK[i], nrow(new.average)), 
                               enVar = rep(unique.params$enVar[i], nrow(new.average)), 
                               mu_base = rep(unique.params$mu_base[i], nrow(new.average)))
    
    df.simParamsAll <- data.frame(mu_inv = rep(unique.params$mu_inv[i], nrow(df.average)), 
                               mig = rep(unique.params$mig1[i], nrow(df.average)),
                               alpha = rep(unique.params$alpha[i], nrow(df.average)), 
                               sigmaK = rep(unique.params$sigmaK[i], nrow(df.average)), 
                               enVar = rep(unique.params$enVar[i], nrow(df.average)), 
                               mu_base = rep(unique.params$mu_base[i], nrow(df.average)))
   
    # create seed columns so we can keep track of relevant seed names
    df.Seedcolumns <- NULL
    df.SeedcolumnsAll <- NULL
    vect.colNames <- NULL
    for(m in 1:reps){
      seedCol <- paste("Seed", m, sep = "")
      if(!is.na(av.seeds[m])){
        df.Seedcolumns <- cbind(df.Seedcolumns, rep(av.seeds[m], nrow(new.average)))
        df.SeedcolumnsAll <- cbind(df.SeedcolumnsAll, rep(av.seeds[m], nrow(df.average)))
      } else {
        df.Seedcolumns <- cbind(df.Seedcolumns, rep(NA, nrow(new.average)))
        df.SeedcolumnsAll <- cbind(df.SeedcolumnsAll, rep(NA, nrow(df.average)))
      }
      vect.colNames <- c(vect.colNames, seedCol)
    }
    colnames(df.Seedcolumns) <- vect.colNames
    colnames(df.SeedcolumnsAll) <- vect.colNames
    
    # finally bind together all the data in the final dataframe
    df.LA.average <- rbind(df.LA.average, cbind(new.average, df.simParams, df.Seedcolumns))
    df.LA.stderr <- rbind(df.LA.stderr, cbind(new.se, df.simParams, df.Seedcolumns))
    df.LA.all <- rbind(df.LA.all, cbind(df.average, df.simParamsAll, df.SeedcolumnsAll))
    count <- count + 1
    print(count)
    print(proc.time() - ptm)
} # close average for loop
###################################################################
df.LA.average <- subset(df.LA.average, select = -seed)
df.LA.stderr <- subset(df.LA.stderr, select = -seed)
df.LA.all <- subset(df.LA.all, select = -seed)

#df.LA.average <- read.table("FullSet_LAaverage.txt", header = TRUE)
write.table(df.LA.average, "FullSet_LAaverage.txt", row.names = FALSE)
write.table(df.LA.stderr, "FullSet_LAstderr.txt", row.names = FALSE)
write.table(df.LA.all, "FullSet_LAallData.txt", row.names = FALSE)

# this should work I don't understand why it doesn't
#df.LA.average[,df.LA.average] <- lapply(df.LA.average[,LA.cols], function(x) as.factor(as.character(x)))
df.LA.average <- subset(df.LA.average, select = -seed)

# convert every column of parameters data to factor 
for(i in 16:ncol(df.LA.average)){
  df.LA.average[,i] <- as.factor(as.character(df.LA.average[,i]))
}
```

# Run this chunk if average file has already been made
```{r}
df.LA.average <- read.table("src/results/FullSet_LAaverage.txt", header = TRUE, stringsAsFactors = FALSE)

for(i in 16:ncol(df.LA.average)){
  df.LA.average[,i] <- as.factor(as.character(df.LA.average[,i]))
}
```

## Plotting

### Plotting for Local Adaptation 
```{r}
library(ggplot2)
inv.labs <- c("0" = "No Inversions", "0.001" = "Inversion Mutation Rate = 0.001")
ggplot(data = df.LA.average[df.LA.average$sim_gen > 4000,], 
       aes(x = sim_gen, y = localAdaptSA, group = interaction(mig, mu_base))) + 
    geom_line(aes(color = mig, linetype = mu_base), size = 0.75) + 
    facet_wrap(~ mu_inv, labeller = labeller(mu_inv = inv.labs)) + 
    labs(title = "Effect of Inversions on Local Adaptation",
           y = "Local Adaptation",
           x = "Generation") +
 # scale_fill_discrete(names = c("Migration", "QTN Mutation Rate")) +
  guides(color = guide_legend(title = "Migration Rate"), 
         linetype = guide_legend(title = "QTN Mutation Rate")) +
  theme_classic() +
  theme(panel.background = element_blank(), 
        strip.background = element_rect(colour = "white", fill = "grey92")) +
  scale_color_manual(values=c("cadetblue1","cadetblue3", "cornflowerblue", "navy")) +
  scale_x_continuous(expand = c(0, 0), limits = c(4200, NA)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1))

```

## Plotting for inversion characteristics
```{r}
#############################################################################################################
#### PLOTTING For inversion characteristics ####
#############################################################################################################
  
  ## make labels for facet wrapping
    alpha.labels <- c("0.2" = "sigmaMu = 0.2", "0.002" = "sigmaMu = 0.002")
    mig.labels <- c("0.001" = "mig = 0.001", "0.01" = "mig = 0.01", "0.1" = "mig = 0.1", 
                  "0.25" = "mig = 0.25", "0.4" = "mig = 0.4", "0.5" = "mig = 0.5")


  ## plot inversion age --  split by inversiion FST with reps
    ggplot(data = df.FSTsplit[df.FSTsplit$enVar == 0 & df.FSTsplit$muInv == 0.001 & 
                                df.FSTsplit$sigmaK == 0.75 & df.FSTsplit$alpha == 0.002,], 
           aes(x = sim_gen, y = inv_ageT10, group = interaction(muBase, rep))) + 
           geom_line(aes(color = muBase), size = 0.75, alpha = 0.3) + 
           geom_line(data = df.aveFSTsplit[df.aveFSTsplit$enVar == 0 & df.aveFSTsplit$muInv == 0.001 & 
                                        df.aveFSTsplit$sigmaK == 0.75 & df.aveFSTsplit$alpha == 0.002,], 
                     aes(x = sim_gen, y = inv_ageT10, group = muBase, color = muBase), size = 1.2) + 
           facet_wrap(~ mig1 , labeller = labeller(mig1 = mig.labels), ncol = 6, nrow = 1) + 
           labs(title = "Average Inversions Age Through Time",
                 y = "Average Inversion Age",
                 x = "Generation") +
           guides(color = guide_legend(title = "QTN Mutation Rate")) +
           theme_classic() +
           theme(panel.background = element_blank(), 
                 strip.background = element_rect(colour = "white", fill = "grey92")) +
           scale_color_manual(values=c( "cadetblue3", "navy")) +
           scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
           scale_y_continuous(expand = c(0, 0), limits = c(0, max(df.FSTsplit$inv_ageT10)))
    
  ## plot inversion age -- split by inversion FST just average
    ggplot(data = df.aveFSTsplit[df.aveFSTsplit$enVar == 0 & df.aveFSTsplit$muInv == 0.001 & 
                              df.aveFSTsplit$sigmaK == 0.75 & df.aveFSTsplit$alpha == 0.002,], 
           aes(x = sim_gen, y = inv_ageT10)) + 
      geom_line(aes(color = mig1), size = 0.75, alpha = 0.95) + 
      geom_line(data = df.aveFSTsplit[df.aveFSTsplit$enVar == 0 & df.aveFSTsplit$muInv == 0.001 & 
                                      df.aveFSTsplit$sigmaK == 0.75 & df.aveFSTsplit$alpha == 0.002,], 
                 aes(x = sim_gen, y = inv_ageB90, color = mig1), 
                 size = 0.75, alpha = 0.95, linetype = 3) + 
      facet_wrap(~ muBase + mig1 , labeller = labeller(mig1 = mig.labels), ncol = 6, nrow = 2) + 
      labs(title = "Average Inversions Age Through Time",
           y = "Average Inversion Age",
           x = "Generation") +
      guides(color = guide_legend(title = "Migration Rate")) +
      scale_color_manual(values=c( "darkgrey","cadetblue1","cadetblue3", 
                                   "cornflowerblue", "navy", "black")) +
      theme_classic() +
      theme(panel.background = element_blank(), 
            strip.background = element_rect(colour = "white", fill = "grey92")) +
      scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
      scale_y_continuous(expand = c(0, 0), limits = c(0, max(df.FSTsplit$inv_ageT10)))
    
    ggplot(data = df.aveFSTsplit[df.aveFSTsplit$enVar == 0 & df.aveFSTsplit$muInv == 0.001 & 
                                   df.aveFSTsplit$sigmaK == 0.75 & df.aveFSTsplit$alpha == 0.002,], 
           aes(x = sim_gen, y = inv_ageT10)) + 
      geom_line(aes(color = muBase), size = 0.75, alpha = 0.95) + 
      geom_line(data = df.aveFSTsplit[df.aveFSTsplit$enVar == 0 & df.aveFSTsplit$muInv == 0.001 & 
                                        df.aveFSTsplit$sigmaK == 0.75 & df.aveFSTsplit$alpha == 0.002,], 
                aes(x = sim_gen, y = inv_ageB90, color = muBase), 
                size = 0.75, alpha = 0.95, linetype = 3) + 
      facet_wrap(~ muBase + mig1 , labeller = labeller(mig1 = mig.labels), ncol = 6, nrow = 2) + 
      labs(title = "Average Inversions Age Through Time",
           y = "Average Inversion Age",
           x = "Generation") +
      guides(color= guide_legend(title = "QTN Mutation Rate")) +
      scale_color_manual(values=c( "cadetblue3", "dodgerblue3")) +
      theme_classic() +
      theme(panel.background = element_blank(), 
            strip.background = element_rect(colour = "white", fill = "grey92")) +
      scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
      scale_y_continuous(expand = c(0, 0), limits = c(0, max(df.aveFSTsplit$inv_ageT10)))
   
  ## plot all data         
    inv.Age <-  ggplot(data = df.all.data[df.all.data$enVar == 0 & df.all.data$muInv == 0.001 &
                                          df.all.data$sigmaK == 0.75 & df.all.data$alpha == 0.002,], 
                       aes(x = sim_gen, y = aveAge, group = interaction(muBase, rep))) + 
      geom_line(aes(color = muBase), size = 0.75, alpha = 0.3) + 
      geom_line(data = df.ave.data[df.ave.data$enVar == 0 & df.ave.data$muInv == 0.001 & 
                                   df.ave.data$sigmaK == 0.75 & df.ave.data$alpha == 0.002,], 
                aes(x = sim_gen, y = aveAge, group = muBase, color = muBase), 
              size = 1.2) + 
      facet_wrap(~ mig1 , labeller = labeller(mig1 = mig.labels),
                 ncol = 6, nrow = 1) + 
      labs(title = "Average Inversions Age Through Time",
           y = "Average Inversion Age",
           x = "Generation") +
      guides(color = guide_legend(title = "QTN Mutation Rate")) +
      theme_classic() +
      theme(panel.background = element_blank(), 
            strip.background = element_rect(colour = "white", fill = "grey92")) +
      scale_color_manual(values=c( "cadetblue3", "navy")) +
      scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
      scale_y_continuous(expand = c(0, 0), limits = c(0, max(df.ave.data$aveAge)))
    
 ### Inversion Length ####
 #######################################
    
    inv.Length <- ggplot(data = df.all.data[df.all.data$enVar == 0 & df.all.data$muInv == 0.001 &
                                               df.all.data$sigmaK == 0.75 & df.all.data$alpha == 0.002,], 
                          aes(x = sim_gen, y = aveLength, group = interaction(muBase, rep))) + 
      geom_line(aes(color = muBase), size = 0.75, alpha = 0.3) + 
      geom_line(data = df.ave.data[df.ave.data$enVar == 0 & df.ave.data$muInv == 0.001 &
                                     df.ave.data$sigmaK == 0.75 & df.ave.data$alpha == 0.002,], 
                aes(x = sim_gen, y = aveLength, group = interaction(muBase), color = muBase), size = 1.2) + 
      facet_wrap(~ mig1 , labeller = labeller(mig1 = mig.labels),
                 ncol = 6, nrow = 1) + 
      labs(title = "Average Inversion Length Through Time",
           y = "Average Inversion Length",
           x = "Generation") +
      guides(color = guide_legend(title = "QTN Mutation Rate")) +
          #   linetype = guide_legend(title = "Sigma Mu (Effect Size)")) +
      theme_classic() +
      theme(panel.background = element_blank(), 
            strip.background = element_rect(colour = "white", fill = "grey92")) +
      scale_color_manual(values=c( "cadetblue3", "navy")) +
      scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
      scale_y_continuous(expand = c(0, 0), limits = c(0, max(df.ave.data$aveLength)))
    
    ## plot inversion length -- split by inversion FST just average
    ggplot(data = df.aveFSTsplit[df.aveFSTsplit$enVar == 0 & df.aveFSTsplit$muInv == 0.001 & 
                                   df.aveFSTsplit$sigmaK == 0.75 & df.aveFSTsplit$alpha == 0.002,], 
           aes(x = sim_gen, y = inv_lengthT10)) + 
      geom_line(aes(color = mig1), size = 0.75, alpha = 0.95) + 
      geom_line(data = df.aveFSTsplit[df.aveFSTsplit$enVar == 0 & df.aveFSTsplit$muInv == 0.001 & 
                                        df.aveFSTsplit$sigmaK == 0.75 & df.aveFSTsplit$alpha == 0.002,], 
                aes(x = sim_gen, y = inv_lengthB90, color = mig1), 
                size = 0.75, alpha = 0.95, linetype = 3) + 
      facet_wrap(~muBase + mig1, labeller = labeller(mig1 = mig.labels), ncol = 6, nrow = 2) + 
      labs(title = "Average Inversions Length Through Time",
           y = "Average Inversion Length",
           x = "Generation") +
      guides(color = guide_legend(title = "Migration Rate")) +
      scale_color_manual(values=c( "darkgrey","cadetblue1","cadetblue3", 
                                   "cornflowerblue", "navy", "black")) +
      theme_classic() +
      theme(panel.background = element_blank(), 
            strip.background = element_rect(colour = "white", fill = "grey92")) +
      scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
      scale_y_continuous(expand = c(0, 0), limits = c(0, max(df.FSTsplit$inv_lengthT10)))
    
    
    ggplot(data = df.aveFSTsplit[df.aveFSTsplit$enVar == 0 & df.aveFSTsplit$muInv == 0.001 & 
                                   df.aveFSTsplit$sigmaK == 0.75 & df.aveFSTsplit$alpha == 0.002,], 
           aes(x = sim_gen, y = num_qtnsT10)) + 
      geom_line(aes(color = mig1), size = 0.75, alpha = 0.95) + 
      geom_line(data = df.aveFSTsplit[df.aveFSTsplit$enVar == 0 & df.aveFSTsplit$muInv == 0.001 & 
                                        df.aveFSTsplit$sigmaK == 0.75 & df.aveFSTsplit$alpha == 0.002,], 
                aes(x = sim_gen, y = numqtnsB90, color = mig1), 
                size = 0.75, alpha = 0.95, linetype = 3) + 
      facet_wrap(~muBase + mig1, labeller = labeller(mig1 = mig.labels), ncol = 6, nrow = 2) + 
      labs(title = "Average Number of Inversion QTNs Through Time",
           y = "Average Number of Inversion QTNs",
           x = "Generation") +
      guides(color = guide_legend(title = "Migration Rate")) +
      scale_color_manual(values=c( "darkgrey","cadetblue1","cadetblue3", 
                                   "cornflowerblue", "navy", "black")) +
      theme_classic() +
      theme(panel.background = element_blank(), 
            strip.background = element_rect(colour = "white", fill = "grey92")) +
      scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
      scale_y_continuous(expand = c(0, 0), limits = c(0, 300))
   
### Inversion QTNs ####
#######################################   
    
inv.numQTNs <-  ggplot(data = df.all.data[df.all.data$enVar == 0 & df.all.data$muInv == 0.001,], 
                          aes(x = sim_gen, y = aveQTNs, group = interaction(muBase, alpha, rep))) + 
      geom_line(aes(color = muBase, linetype = alpha), size = 0.75, alpha = 0.3) + 
      geom_line(data = df.ave.data[df.ave.data$enVar == 0 & df.ave.data$muInv == 0.001,], 
                aes(x = sim_gen, y = aveQTNs, group = interaction(muBase, alpha), linetype = alpha, color = muBase), size = 1.2) + 
      facet_wrap(~ sigmaK + mig1 , labeller = labeller(mig1 = mig.labels),
                 ncol = 6, nrow = 3) + 
      labs(title = "Average Number of Inversion QTNs Through Time",
           y = "Average Number of Inversion QTNs",
           x = "Generation") +
      guides(color = guide_legend(title = "QTN Mutation Rate"), 
             linetype = guide_legend(title = "Sigma Mu (Effect Size)")) +
      theme_classic() +
      theme(panel.background = element_blank(), 
            strip.background = element_rect(colour = "white", fill = "grey92")) +
      scale_color_manual(values=c( "cadetblue3", "navy")) +
      scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
      scale_y_continuous(expand = c(0, 0), limits = c(0, max(df.ave.data$aveQTNs)))
    


```
## Dataframe for Katie

```{r}
# merge inversion through time data with parameters
df.invTimeParam <- merge(df.invTime, df.params, all.x = TRUE, by = c("seed"))

# subset inversion qtn file for just the seeds of interest
df.invQTNTimeSub <- subset(df.invQTNTime, subset = seed == c(3383645, 3383657, 3383669, 3383681, 3383693))
colnames(df.invQTNData)[3] <- "qtn_id"

# merge inversion QTN data with parameters
movie.InvQTNSub.Data <- merge(df.invQTNTimeSub, df.params, all.x = TRUE, by = c("seed"))
head(movie.InvQTNSub.Data)

df.invQTNData[df.invQTNData$sim_gen==200 & df.invQTNData$qtn_id == 9084,]
movie.InvQTN.Data <- merge(df.invQTNData, movie.InvQTNSub.Data, all.y = TRUE, by = c("seed", "qtn_id"))
movie.Inv.Data <- subset(df.invTimeParam, subset = muBase == 0.0000001 
                          & muInv == 0.001 & sigmaK == 0.75 & alpha == 0.002 
                          & enVar == 0 & mig1 == 0.25)
movie.Inv.Data <- merge(movie.Inv.Data, all.y = TRUE, by = c("seed", "inv_id"))


write.table(movie.Inv.Data, "movieInvData.txt", row.names = FALSE)
write.table(movie.InvQTN.Data, "movieInvQTNData.txt", row.names = FALSE)

```


# Subset for a file that you are interested in
```{r}
# Run subset for either the seed you want or the parameter values
# either this line:
df.singleFile <- subset(df.params, subset = seed == 3383282)
# or this line:
df.singleFile <- subset(df.params, subset = muBase == 1e-7 & muInv == 1e-3 & alpha == 0.002 & sigmaK == 0.75 & 
                                            rep == 1 & enVar == 0, mig1 == 0.25)

```