---
title: "processInversionSims"
author: "Sara Michele Schaal"
date: "6/17/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
```

```{r}
options(scipen = 999)
### Read in datafiles ###
#df.params <- read.table("fullINVparamslist.txt", header = TRUE)
#df.params <- read.table("src/testINVparamslist.txt")
#df.params <- data.frame(seed = c(1751625359138, 1751583067058, 1751620201657, 1751583946190, 1751616892960,
                              #   1751566183207, 1751570485992, 1751568804886, 1751626452845, 2065972841416,
                              #   1751582301417, 1751629252645, 1751629716441, 1751572690982, 1751628536499, 
                              #   1751630159938, 1614479103543, 1614479625159, 1614479993467))

df.params <- read.table("invSimParams.txt", header = TRUE)
```


```{r}
folder <- "../results/"

## Inversion Through Time
df.invTime <- NULL
count <- 0
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  invTimeNewFile <- read.table(paste(folder, seed, "_outputInvTime.txt", sep=""), header = TRUE, stringsAsFactors = FALSE)
  if(nrow(invTimeNewFile) > 0){
    invTimeNewFile$seed <- seed
    df.invTime <-  rbind(df.invTime, invTimeNewFile)
  }
  count <- count + 1
  print(count)
}
write.table(df.invTime, "FullSet_invTime.txt", row.names = FALSE)

## Inversion Summary Data
df.invData <- NULL
count <- 0
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  invData <- read.table(paste(folder, seed, "_outputInvSumInfo.txt", sep=""), header = TRUE,
                        stringsAsFactors = FALSE)
  if(nrow(invData) > 0){
    invData$seed <- seed
    df.invData <- rbind(df.invData, invData)
  }
  count <- count + 1
  print(count)
}
write.table(df.invData, "FullSet_invData.txt", row.names = FALSE)

## QTN Inversion Through Time
df.invQTNTime <- NULL
count <- 0
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  invQTNTimeNewFile <- read.table(paste(folder, seed, "_outputInvQtn.txt", sep=""), header = TRUE, stringsAsFactors = FALSE)
  if(nrow(invQTNTimeNewFile) > 0){
    invQTNTimeNewFile$seed <- seed
    df.invQTNTime <-  rbind(df.invQTNTime, invQTNTimeNewFile)
  }
  count <- count + 1
  print(count)
}
write.table(df.invQTNTime, "FullSet_invQTNTime.txt", row.names = FALSE)

## QTN Inversion Summary Data
df.invQTNData <- NULL
count <- 0
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  invQTNDataNewFile <- read.table(paste(folder, seed, "_outputInvQtnSumInfo.txt", sep=""), header = TRUE, stringsAsFactors = FALSE)
  if(nrow(invQTNDataNewFile) > 0){
    invQTNDataNewFile$seed <- seed
    df.invQTNData <-  rbind(df.invQTNData, invQTNDataNewFile)
  }
  count <- count + 1
  print(count)
}
write.table(df.invQTNData, "FullSet_invQTNSumInfo.txt", row.names = FALSE)

## Pop Dynamics
df.popDyn <- NULL
count <- 0
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  popDynNewFile <- read.table(paste(folder, seed, "_outputPopDynam.txt", sep=""), header = TRUE, stringsAsFactors = FALSE)
  if(nrow(popDynNewFile) > 0){
   popDynNewFile$seed <- seed
   df.popDyn <-  rbind(df.popDyn, popDynNewFile)
  }
  count <- count + 1
  print(count)
}
write.table(df.popDyn, "FullSet_popDyn.txt", row.names = FALSE)

#SimStats
df.simStats <- NULL
count <- 0
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  simStatsNewFile <- read.table(paste(folder, seed, "_outputSimStats.txt", sep=""), stringsAsFactors = FALSE)
  simStatsNewFile$seed <- seed
  df.simStats <-  rbind(df.simStats, simStatsNewFile)
  count <- count + 1
  print(count)
}
df.simStats <- df.simStats[,2:ncol(df.simStats)]
colnames(df.simStats) <- c("mig1", "mig2", "pop1N", "pop2N", "mu_base", "mu_inv", "r", "alpha", "sigmaK", "burnin", "dom", "enVar", "Seed")
write.table(df.simStats, "FullSet_simStats.txt", row.names = FALSE)

#Mutations
df.finalMuts <- NULL
count <- 0
no.Data <- NULL
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  if(file.exists(paste(folder, seed, "_outputMutations.txt", sep=""))){
    finalMutsNewFile <- read.table(paste(folder, seed, "_outputMutations.txt", sep=""), stringsAsFactors = FALSE, header = TRUE)
  finalMutsNewFile$seed <- seed
  df.finalMuts <-  rbind(df.finalMuts, finalMutsNewFile)
  } else {
    no.Data <- c(no.Data, seed)
  }
  count <- count + 1
  print(count)
}
write.table(no.Data,"seedsToRerun.txt", row.names = FALSE)
colnames(df.finalMuts)[2] <- "mut_id" 
file.exists(paste(folder, seed, "_outputMutations.txt", sep=""))
write.table(df.finalMuts, "FullSet_finalMuts.txt", row.names = FALSE)

```

## Combine dataframes

```{r}
df.mutInfo <- merge(df.finalMuts, df.invQTNData, all.x = TRUE, by = c("seed", "mut_id"))
df.mutInfo <- merge(df.mutInfo, df.params, all.x = TRUE, by = c("seed"))
df.invMutInfo <- subset(df.mutInfo, subset = muInv > 0)
df.invMutInfo$inOut <- ifelse(is.na(df.invMutInfo$inv_id), "out", "in")

mig.labels <- c("0.001" = "mig = 0.001", "0.01" = "mig = 0.01", "0.1" = "mig = 0.1", 
                  "0.25" = "mig = 0.25", "0.4" = "mig = 0.4", "0.5" = "mig = 0.5")
levels(df.invMutInfo$muBase) <- c("1e-7", "1e-6")

 for(i in 18:ncol(df.invMutInfo)){
    df.invMutInfo[,i] <- as.factor(as.character(df.invMutInfo[,i]))
  }  
  
ggplot(data = df.invMutInfo[df.invMutInfo$enVar == 0 & df.invMutInfo$muInv == 0.001 & df.invMutInfo$alpha == 0.002 & df.invMutInfo$muBase == "1e-6",], 
       aes(x = sigmaK, y = adapPhenoDiv, fill = inOut)) + 
  geom_boxplot(outlier.colour="black", outlier.shape=16,
             outlier.size=1, notch=FALSE) +
  facet_wrap(~mig1, labeller = labeller(mig1 = mig.labels), ncol = 6) +
 scale_y_continuous(expand = c(0, 0), limits = c(-0.008, 0.008))

```

## Dataframe for Katie

```{r}
# merge inversion through time data with parameters
df.invTimeParam <- merge(df.invTime, df.params, all.x = TRUE, by = c("seed"))

# subset inversion qtn file for just the seeds of interest
df.invQTNTimeSub <- subset(df.invQTNTime, subset = seed == c(3383645, 3383657, 3383669, 3383681, 3383693))
colnames(df.invQTNData)[3] <- "qtn_id"

# merge inversion QTN data with parameters
movie.InvQTNSub.Data <- merge(df.invQTNTimeSub, df.params, all.x = TRUE, by = c("seed"))
head(movie.InvQTNSub.Data)

df.invQTNData[df.invQTNData$sim_gen==200 & df.invQTNData$qtn_id == 9084,]
movie.InvQTN.Data <- merge(df.invQTNData, movie.InvQTNSub.Data, all.y = TRUE, by = c("seed", "qtn_id"))
movie.Inv.Data <- subset(df.invTimeParam, subset = muBase == 0.0000001 
                          & muInv == 0.001 & sigmaK == 0.75 & alpha == 0.002 
                          & enVar == 0 & mig1 == 0.25)
movie.Inv.Data <- merge(movie.Inv.Data, all.y = TRUE, by = c("seed", "inv_id"))


write.table(movie.Inv.Data, "movieInvData.txt", row.names = FALSE)
write.table(movie.InvQTN.Data, "movieInvQTNData.txt", row.names = FALSE)

```
## Unique Parameters - (used for averaging)
Subset original parameters dataframe to identify all unique parameter combinations. Then using those parameter values identify the simulation seeds of replicate simulations for each set of parameters.🤯
```{r}

# find all unique parameter combinations 
unique.params <- unique(df.simStats[c("mig1","mig2", "pop1N", "pop2N", "mu_base", "mu_inv", "r", "alpha", "sigmaK", "burnin", "dom", "enVar")])
reps <- 5
# make columns for replicate seeds
for(i in 1:reps){
  seedColName <- paste("Seed", i, sep = "")
  unique.params[,seedColName] <- NA
}

count <- 0
for(i in 1:nrow(unique.params)){
  for(j in 1:nrow(df.simStats)){
    if(apply(unique.params[i,1:12], 1, paste, collapse = " ") == 
       apply(df.simStats[j,c("mig1","mig2", "pop1N", "pop2N", "mu_base", "mu_inv", "r", "alpha", 
                             "sigmaK", "burnin", "dom", "enVar")], 1, paste, collapse = " ")){
      for(k in 1:reps){
        seedCol <- paste("Seed", k, sep="")
        if(is.na(unique.params[i, seedCol])){
          unique.params[i, seedCol] <- df.simStats$Seed[j]
          break
        } 
      }
    }
  }
  count <- count + 1
  print(count)
}
write.table(unique.params, "FullSet_uniqueParams", row.names = FALSE)

```

### Average Replicate Simulations - Pop dynamics file
Create a dataframe that has the average of all the replicate parameter values. Then bind the corresponding parameters to the dataframe for ease of plotting and analysis later. 
```{r}
##################################
### convert all "NAN"s to NAs  ### REMOVE WITH NEW SIMs CHANGING TO ZEROS
#########################################################################

for(i in 1:nrow(df.popDyn)){
  for(j in 1:ncol(df.popDyn)){
    if(df.popDyn[i,j] == "NAN"){
      df.popDyn[i,j] <- 0
    }
  }
}
# convert every column to numeric ? might still need
cols <- c(1:ncol(df.popDyn))
df.popDyn[, cols] <- apply(df.popDyn[,cols], 2, as.numeric)
###################################################################


########################
### average for loop ###
###################################################################

df.LA.average <- NULL
df.LA.stderr <- NULL
df.LA.all <- NULL
count <- 0
# step through each line of the unique parameters dataframe
for(i in 1:nrow(unique.params)){
    ptm <- proc.time()
    # create empty variable for putting each data set that should be average
    df.average <- NULL
    av.seeds <- NULL
    # step through the population dynamics dataframe
    for(j in 1:nrow(df.popDyn)){
      # step through the different seeds that have the unique parameters
      for(k in 1:reps){
        seedCol <- paste("Seed", k, sep="")
        # if that seed is not an NA then do the next step
        if(!is.na(unique.params[i, seedCol])){
          # if the seed from unique parameters matches the seed of the pop dynam dataset store it
          if(df.popDyn$seed[j] == unique.params[i, seedCol]){
            # bind data from same parameter combination together for averaging later
            df.average <- rbind(df.average, df.popDyn[j,])
            # store unique seeds to this dataset
            av.seeds <- unique(c(av.seeds, unique.params[i, seedCol]))
          }
        }
      }
    }
    
    
    ## average columns of interest
    # df.averageSubset <- df.average[,vect.aggCols]
    new.average <- aggregate(.~sim_gen, data = df.average, FUN = mean)
    SE <- function(x){
      sd(x)/sqrt(length(x))
    }
    new.se <- aggregate(.~sim_gen, data = df.average, FUN = SE)
    df.simParams <- data.frame(mu_inv = rep(unique.params$mu_inv[i], nrow(new.average)), 
                               mig = rep(unique.params$mig1[i], nrow(new.average)),
                               alpha = rep(unique.params$alpha[i], nrow(new.average)), 
                               sigmaK = rep(unique.params$sigmaK[i], nrow(new.average)), 
                               enVar = rep(unique.params$enVar[i], nrow(new.average)), 
                               mu_base = rep(unique.params$mu_base[i], nrow(new.average)))
    
    df.simParamsAll <- data.frame(mu_inv = rep(unique.params$mu_inv[i], nrow(df.average)), 
                               mig = rep(unique.params$mig1[i], nrow(df.average)),
                               alpha = rep(unique.params$alpha[i], nrow(df.average)), 
                               sigmaK = rep(unique.params$sigmaK[i], nrow(df.average)), 
                               enVar = rep(unique.params$enVar[i], nrow(df.average)), 
                               mu_base = rep(unique.params$mu_base[i], nrow(df.average)))
   
    # create seed columns so we can keep track of relevant seed names
    df.Seedcolumns <- NULL
    df.SeedcolumnsAll <- NULL
    vect.colNames <- NULL
    for(m in 1:reps){
      seedCol <- paste("Seed", m, sep = "")
      if(!is.na(av.seeds[m])){
        df.Seedcolumns <- cbind(df.Seedcolumns, rep(av.seeds[m], nrow(new.average)))
        df.SeedcolumnsAll <- cbind(df.SeedcolumnsAll, rep(av.seeds[m], nrow(df.average)))
      } else {
        df.Seedcolumns <- cbind(df.Seedcolumns, rep(NA, nrow(new.average)))
        df.SeedcolumnsAll <- cbind(df.SeedcolumnsAll, rep(NA, nrow(df.average)))
      }
      vect.colNames <- c(vect.colNames, seedCol)
    }
    colnames(df.Seedcolumns) <- vect.colNames
    colnames(df.SeedcolumnsAll) <- vect.colNames
    
    # finally bind together all the data in the final dataframe
    df.LA.average <- rbind(df.LA.average, cbind(new.average, df.simParams, df.Seedcolumns))
    df.LA.stderr <- rbind(df.LA.stderr, cbind(new.se, df.simParams, df.Seedcolumns))
    df.LA.all <- rbind(df.LA.all, cbind(df.average, df.simParamsAll, df.SeedcolumnsAll))
    count <- count + 1
    print(count)
    print(proc.time() - ptm)
} # close average for loop
###################################################################
df.LA.average <- subset(df.LA.average, select = -seed)
df.LA.stderr <- subset(df.LA.stderr, select = -seed)
df.LA.all <- subset(df.LA.all, select = -seed)

#df.LA.average <- read.table("FullSet_LAaverage.txt", header = TRUE)
write.table(df.LA.average, "FullSet_LAaverage.txt", row.names = FALSE)
write.table(df.LA.stderr, "FullSet_LAstderr.txt", row.names = FALSE)
write.table(df.LA.all, "FullSet_LAallData.txt", row.names = FALSE)

df.LA.average <- read.table(paste0(folder, "FullSet_LAaverage.txt"), header = TRUE, stringsAsFactors = FALSE)

# df.LA.average[,df.LA.average] <- lapply(df.LA.average[,LA.cols], function(x) as.factor(as.character(x)))
# convert every column of parameters data to factor 

for(i in 16:ncol(df.LA.average)){
  df.LA.average[,i] <- as.factor(as.character(df.LA.average[,i]))
}


```

### Plot Local Adaptation 
```{r}
library(ggplot2)
inv.labs <- c("0" = "No Inversions", "1e-06" = "Moderate Inversion Mu", "0.001" = "High Inversion Mu")
df.LA.average$mu_inv <- factor(df.LA.average$mu_inv, levels = c("0","1e-06", "0.001"))
df.LA.average$mu_base <- factor(df.LA.average$mu_base, levels = c("high", "low"))
high.alpha <- subset(df.LA.average, subset = alpha == 0.2 & enVar == 0 )
low.alpha <- subset(df.LA.average, subset = alpha == 0.002 & enVar == 0 )

### Panels are inversion vs. no inversion ###
highSigMu.LAinv <- ggplot(data = high.alpha, 
       aes(x = sim_gen, y = localAdaptSA, group = interaction(mig, mu_base))) + 
       geom_line(aes(color = mig, linetype = mu_base), size = 0.75) + 
       facet_wrap(~ sigmaK + mu_inv, labeller = labeller(mu_inv = inv.labs)) + 
       labs(title = "Local Adaptation - high sigma_mu (0.2)",
            y = "Local Adaptation",
            x = "Generation") +
 #scale_fill_discrete(names = c("Migration", "QTN Mutation Rate")) +
       guides(color = guide_legend(title = "Migration Rate"),
              linetype = guide_legend(title = "QTN Mutation Rate")) +
       theme_classic() +
       theme(panel.background = element_blank(), 
             strip.background = element_rect(colour = "white", fill = "grey92")) +
#  theme(legend.position = "none") +
       scale_linetype_discrete(name = "QTN Mutation Rate", labels= c("1e-6", "1e-7")) +
       scale_color_manual(values=c("darkgrey","cadetblue1","cadetblue3",
                                   "cornflowerblue", "navy","black")) +
       scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
       scale_y_continuous(expand = c(0, 0), limits = c(-0.1, 1))

lowSigMu.LAinv <- ggplot(data = low.alpha, 
       aes(x = sim_gen, y = localAdaptSA, group = mig)) + 
       geom_line(aes(color = mig), size = 0.75) + 
       facet_wrap(~sigmaK + mu_inv, labeller = labeller(mu_inv = inv.labs)) + 
       labs(title = "Local Adaptation - low sigma_mu (0.002)",
            y = "Local Adaptation",
            x = "Generation") +
 # scale_fill_discrete(names = c("Migration", "QTN Mutation Rate")) +
       guides(color = guide_legend(title = "Migration Rate")) +
             # linetype = guide_legend(title = "QTN Mutation Rate")) +
       theme_classic() +
       theme(panel.background = element_blank(), 
             strip.background = element_rect(colour = "white", fill = "grey92")) +
#  theme(legend.position = "none") +
       scale_color_manual(values=c("darkgrey","cadetblue1","cadetblue3", 
                                   "cornflowerblue", "navy", "black")) +
       scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
       scale_y_continuous(expand = c(0, 0), limits = c(-0.1, 1))


```

```{r}
library(dplyr)

df.LA.average %>%
  filter(alpha == 0.002, sim_gen == 50000, mu_inv == 0.001, sigmaK == 0.75, enVar == 0) 



```