---
title: "processInversionSims"
author: "Sara Michele Schaal"
date: "6/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
options(scipen = 999)
### Read in datafiles ###
############################################
#df.params <- read.table("fullINVparamslist.txt", header = TRUE)
#df.params <- read.table("src/testINVparamslist.txt")
#df.params1 <- read.table("src/testParams.txt", header = TRUE)
#df.params2 <- read.table("src/testTestParams.txt", header = TRUE)
#df.params <- rbind(df.params2, df.params1)

# This searches the specified directory for all files that start with 13 digits (i.e., seeds) and
# only prints those files with unique seeds
df.params <- as.data.frame(unique(substr(list.files("./results/Inversion/20200720_PrelimSetofSims", pattern = "\\d{7}"), 1, 7)))
colnames(df.params) <- "seed"
df.params <- data.frame(seed = 3383332)
not.done.seeds <- c(3383334, 3383330, 3383318, 3383314)
############################################
```


```{r}
folder <- "results/Inversion/20200720_PrelimSetofSims/"
## Inversion Through Time
df.invTime <- NULL
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  invTimeNewFile <- read.table(paste(folder, seed, "_outputInvTime.txt", sep=""), header = TRUE, stringsAsFactors = FALSE)
  if(nrow(invTimeNewFile) > 0){
    invTimeNewFile$seed <- seed
    df.invTime <-  rbind(df.invTime, invTimeNewFile)
  }
}

df.invData <- NULL
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  invData <- read.table(paste(folder, seed, "_outputInvSumInfo.txt", sep=""), header = TRUE,
                        stringsAsFactors = FALSE)
  if(nrow(invData) > 0){
    invData$seed <- seed
    df.invData <- rbind(df.invData, invData)
  }
}

## Pop Dynamics
df.popDyn <- NULL
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  popDynNewFile <- read.table(paste(folder, seed, "_outputPopDynam.txt", sep=""), header = TRUE, stringsAsFactors = FALSE)
  if(nrow(popDynNewFile) > 0){
   popDynNewFile$seed <- seed
   df.popDyn <-  rbind(df.popDyn, popDynNewFile)
  }
}

## Pop Dynamics
df.invQTN <- NULL
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  invQTNsNewFile <- read.table(paste(folder, seed, "_outputInvQtnSumInfo.txt", sep=""), header = TRUE, stringsAsFactors = FALSE)
  if(nrow(invQTNsNewFile) > 0){
   invQTNsNewFile$seed <- seed
   df.invQTN <-  rbind(df.invQTN, invQTNsNewFile)
  }
}


#SimStats
df.simStats <- NULL
for(i in 1:nrow(df.params)){
  seed <- df.params$seed[i]
  if(seed %in% not.done.seeds){
  } else {
    simStatsNewFile <- read.table(paste(folder, seed, "_outputSimStats.txt", sep=""), stringsAsFactors = FALSE)
    simStatsNewFile$seed <- seed
    df.simStats <-  rbind(df.simStats, simStatsNewFile)
  }
}
df.simStats <- df.simStats[,2:ncol(df.simStats)]
colnames(df.simStats) <- c("mig1", "mig2", "pop1N", "pop2N", "mu_base", "mu_inv", "r", "alpha", "sigmaK", "burnin", "dom", "enVar", "Seed")

```

## Unique Parameters
Subset original parameters dataframe to identify all unique parameter combinations. Then using those parameter values identify the simulation seeds of replicate simulations for each set of parameters.
```{r}

unique.params <- unique(df.simStats[c("mig1","mig2", "pop1N", "pop2N", "mu_base", "mu_inv", "r", "alpha", "sigmaK", "burnin", "dom", "enVar")])
reps <- 2
for(i in 1:reps){
  seedColName <- paste("Seed", i, sep = "")
  unique.params[,seedColName] <- NA
}

for(i in 1:nrow(unique.params)){
  for(j in 1:nrow(df.simStats)){
    if(apply(unique.params[i,1:12], 1, paste, collapse = " ") == 
       apply(df.simStats[j,c("mig1","mig2", "pop1N", "pop2N", "mu_base", "mu_inv", "r", "alpha", 
                             "sigmaK", "burnin", "dom", "enVar")], 1, paste, collapse = " ")){
      for(k in 1:reps){
        seedCol <- paste("Seed", k, sep="")
        if(is.na(unique.params[i, seedCol])){
          unique.params[i, seedCol] <- as.character(df.simStats$Seed[j])
          break
        } 
      }
    }
  }
}

```

### Average Replicate Simulations - Pop dynamics file
Create a dataframe that has the average of all the replicate parameter values. Then bind the corresponding parameters to the dataframe for ease of plotting and analysis later. 
```{r}
##################################
### convert all "NAN"s to NAs  ### REMOVE WITH NEW SIMs CHANGING TO ZEROS
#########################################################################

for(i in 1:nrow(df.popDyn)){
  for(j in 1:ncol(df.popDyn)){
    if(df.popDyn[i,j] == "NAN" | df.popDyn[i,j] == "NaN"){
      df.popDyn[i,j] <- 0
    }
  }
}
# convert every column to numeric ? might still need
cols <- c(1:ncol(df.popDyn))
df.popDyn[, cols] <- apply(df.popDyn[,cols], 2, as.numeric)
###################################################################


########################
### average for loop ###
###################################################################

df.LA.average <- NULL
# step through each line of the unique parameters dataframe
for(i in 1:nrow(unique.params)){
    # create empty variable for putting each data set that should be average
    df.average <- NULL
    av.seeds <- NULL
    # step through the population dynamics dataframe
    for(j in 1:nrow(df.popDyn)){
      # step through the different seeds that have the unique parameters
      for(k in 1:reps){
        seedCol <- paste("Seed", k, sep="")
        # if that seed is not an NA then do the next step
        if(!is.na(unique.params[i, seedCol])){
          # if the seed from unique parameters matches the seed of the pop dynam dataset store it
          if(df.popDyn$seed[j] == unique.params[i, seedCol]){
            df.average <- rbind(df.average, df.popDyn[j,])
            av.seeds <- unique(c(av.seeds, unique.params[i, seedCol]))
          }
        }
      }
    }
    
    # ## identify columns with no data -- REMOVE CHANGING TO ZEROS NOT NAs
    # vect.NAlist <- sapply(df.average, function(x)all(!is.na(x)))
    # 
    # vect.aggCols <- NULL
    # vect.noAggCols <- NULL
    # for(p in 1:length(vect.NAlist)){
    #   if(vect.NAlist[p] & names(vect.NAlist[p]) != "seed"){
    #     vect.aggCols <- c(vect.aggCols, names(vect.NAlist)[p])
    #   } else if(names(vect.NAlist[p]) != "seed") {
    #     vect.noAggCols <- c(vect.noAggCols, names(vect.NAlist)[p])
    #   }
    # }
    
    ## average columns of interest
   # df.averageSubset <- df.average[,vect.aggCols]
    new.average <- aggregate(.~sim_gen, data = df.average, FUN = mean)
    df.simParams <- data.frame(mu_inv = rep(unique.params$mu_inv[i], nrow(new.average)), 
                               mig = rep(unique.params$mig1[i], nrow(new.average)),
                               alpha = rep(unique.params$alpha[i], nrow(new.average)), 
                               sigmaK = rep(unique.params$sigmaK[i], nrow(new.average)), 
                               enVar = rep(unique.params$enVar[i], nrow(new.average)), 
                               mu_base = rep(unique.params$mu_base[i], nrow(new.average)))
   
    # ## create NA dataframe for columns with no data -- REMOVE CHANGING TO ZEROS NOT NAs
    # df.NAcols <- matrix(NA, nrow = nrow(new.average), ncol = length(vect.noAggCols))
    # colnames(df.NAcols) <- vect.noAggCols
    
    # create seed columns so we can keep track of relevant seed names
    df.Seedcolumns <- NULL
    vect.colNames <- NULL
    vect.NAseeds <- NULL
    for(m in 1:reps){
      seedCol <- paste("Seed", m, sep = "")
      if(!is.na(av.seeds[m])){
        df.Seedcolumns <- cbind(df.Seedcolumns, rep(av.seeds[m], nrow(new.average)))
      } else {
        df.Seedcolumns <- cbind(df.Seedcolumns, rep(NA, nrow(new.average)))
      }
      vect.colNames <- c(vect.colNames, seedCol)
    }
    colnames(df.Seedcolumns) <- vect.colNames
    
    # finally bind together all the data in the final dataframe
    df.LA.average <- rbind(df.LA.average, cbind(new.average, df.simParams, df.Seedcolumns))
    
} # close average for loop
###################################################################
df.LA.average <- subset(df.LA.average, select = -seed)

# convert every column of parameters data to factor 
for(i in 25:ncol(df.LA.average)){
  df.LA.average[,i] <- as.factor(as.character(df.LA.average[,i]))
}


```

### Plot Local Adaptation 
```{r}
library(ggplot2)
inv.labs <- c("0" = "No Inversions", "0.001" = "Inversion Mu = 0.001")
high.alpha <- subset(df.LA.average, subset = alpha == 0.2)
low.alpha <- subset(df.LA.average, subset = alpha == 0.002)

highSigMu.LA <- ggplot(data = high.alpha, 
       aes(x = sim_gen, y = localAdaptSA, group = interaction(mig, mu_base))) + 
    geom_line(aes(color = mig, linetype = mu_base), size = 0.75) + 
    facet_wrap(~ mu_inv, labeller = labeller(mu_inv = inv.labs)) + 
    labs(title = "Local Adaptation - high sigma_mu (0.2)",
           y = "Local Adaptation",
           x = "Generation") +
 # scale_fill_discrete(names = c("Migration", "QTN Mutation Rate")) +
  guides(color = guide_legend(title = "Migration Rate"), 
         linetype = guide_legend(title = "QTN Mutation Rate")) +
  theme_classic() +
  theme(panel.background = element_blank(), 
        strip.background = element_rect(colour = "white", fill = "grey92")) +
  theme(legend.position = "none") +
  scale_color_manual(values=c("cadetblue1","cadetblue3", "cornflowerblue", "navy")) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(-0.01, 1))

lowSigMu.LA <- ggplot(data = low.alpha, 
       aes(x = sim_gen, y = localAdaptSA, group = interaction(mig, mu_base))) + 
    geom_line(aes(color = mig, linetype = mu_base), size = 0.75) + 
    facet_wrap(~ mu_inv, labeller = labeller(mu_inv = inv.labs)) + 
    labs(title = "Local Adaptation",
           y = "Local Adaptation",
           x = "Generation") +
 # scale_fill_discrete(names = c("Migration", "QTN Mutation Rate")) +
  guides(color = guide_legend(title = "Migration Rate"), 
         linetype = guide_legend(title = "QTN Mutation Rate")) +
  theme_classic() +
  theme(panel.background = element_blank(), 
        strip.background = element_rect(colour = "white", fill = "grey92")) +
  theme(legend.position = "none") +
  scale_color_manual(values=c("cadetblue1","cadetblue3", "cornflowerblue", "navy")) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1))


```
## test 20200903
```{r}
LAtestplot <- ggplot(data = df.LA.average, 
       aes(x = sim_gen, y = localAdaptSA)) +
       geom_line(aes(color = mu_inv, linetype = sigmaK),  size = 0.75) +
       annotate(geom = "text", x = 2500, y = 0.2, label = "mig = 0.4\nalpha = 0.2\nQTNmut = 1e-9\nenvar = 0.1") +
       theme_classic() +
       theme(panel.background = element_blank(), 
        strip.background = element_rect(colour = "white", fill = "grey92"))

```
# Plot Fitness 
```{r}

highSigMu.Fit <- ggplot(data = high.alpha, 
       aes(x = sim_gen, y = meanFitP1, group = interaction(mig, mu_base))) + 
    geom_line(aes(color = mig, linetype = mu_base), size = 0.75) + 
    facet_wrap(~ mu_inv, labeller = labeller(mu_inv = inv.labs)) + 
    labs(title = "Fitness",
           y = "Fitness",
           x = "Generation") +
  guides(color = guide_legend(title = "Migration Rate"), 
         linetype = guide_legend(title = "QTN Mutation Rate")) +
  theme_classic() +
  theme(panel.background = element_blank(), 
        strip.background = element_rect(colour = "white", fill = "grey92")) +
  theme(legend.position = "none") +
  scale_color_manual(values=c("cadetblue1","cadetblue3", "cornflowerblue", "navy")) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1))

lowSigMu.Fit <- ggplot(data = low.alpha, 
       aes(x = sim_gen, y = meanFitP1, group = interaction(mig, mu_base))) + 
    geom_line(aes(color = mig, linetype = mu_base), size = 0.75) + 
    facet_wrap(~ mu_inv, labeller = labeller(mu_inv = inv.labs)) + 
    labs(title = "Fitness",
           y = "Fitness",
           x = "Generation") +
  guides(color = guide_legend(title = "Migration Rate"), 
         linetype = guide_legend(title = "QTN Mutation Rate")) +
  theme_classic() +
  theme(panel.background = element_blank(), 
        strip.background = element_rect(colour = "white", fill = "grey92")) +
  theme(legend.position = "none") +
  scale_color_manual(values=c("cadetblue1","cadetblue3", "cornflowerblue", "navy")) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1))

```
## Plot phenotypes
```{r}
library(gridExtra)
highSigMu.phenoP1 <- ggplot(data = high.alpha, 
       aes(x = sim_gen, y = meanPhenoP1, group = interaction(mig, mu_base))) + 
    geom_line(aes(color = mig, linetype = mu_base), size = 0.75) + 
    facet_wrap(~ mu_inv, labeller = labeller(mu_inv = inv.labs)) + 
    labs(title = "Phenotype - high sigma_mu (0.2)",
           y = "Phenotype (Population 1)",
           x = "Generation") +
  #guides(color = guide_legend(title = "Migration Rate"), 
        # linetype = guide_legend(title = "QTN Mutation Rate")) +
  theme_classic() +
  theme(panel.background = element_blank(), 
        strip.background = element_rect(colour = "white", fill = "grey92")) +
  theme(legend.position = "none") +
  scale_color_manual(values=c("cadetblue1","cadetblue3", "cornflowerblue", "navy")) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(min(high.alpha$meanPhenoP1)-1, max(high.alpha$meanPhenoP1) +1))

highSigMu.phenoP1Zoom <- ggplot(data = high.alpha[high.alpha$sim_gen >= 10200,], 
       aes(x = sim_gen, y = meanPhenoP1, group = interaction(mig, mu_base))) + 
    geom_line(aes(color = mig, linetype = mu_base), size = 0.75) + 
    geom_hline(yintercept=1, color = "firebrick", size = 1) + 
    facet_wrap(~ mu_inv, labeller = labeller(mu_inv = inv.labs)) + 
    labs(title = " ",
         y = " ",
         x = "Generation") +
  guides(color = guide_legend(title = "Migration Rate"), 
         linetype = guide_legend(title = "QTN Mutation Rate")) +
  theme_classic() +
  theme(panel.background = element_blank(), 
        strip.background = element_rect(colour = "white", fill = "grey92")) +
  scale_color_manual(values=c("cadetblue1","cadetblue3", "cornflowerblue", "navy")) +
  scale_x_continuous(expand = c(0, 0), limits = c(10000, NA)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(-1, 2))

lowSigMu.phenoP1 <- ggplot(data = low.alpha, 
       aes(x = sim_gen, y = meanPhenoP1, group = interaction(mig, mu_base))) + 
    geom_line(aes(color = mig, linetype = mu_base), size = 0.75) + 
    facet_wrap(~ mu_inv, labeller = labeller(mu_inv = inv.labs)) + 
    labs(title = "Phenotype - low sigma_mu (0.002)",
           y = "Phenotype (Population 1)",
           x = "Generation") +
  #guides(color = guide_legend(title = "Migration Rate"), 
        # linetype = guide_legend(title = "QTN Mutation Rate")) +
  theme_classic() +
  theme(panel.background = element_blank(), 
        strip.background = element_rect(colour = "white", fill = "grey92")) +
  theme(legend.position = "none") +
  scale_color_manual(values=c("cadetblue1","cadetblue3", "cornflowerblue", "navy")) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(min(low.alpha$meanPhenoP1)-0.5, max(low.alpha$meanPhenoP1) +0.5))

lowSigMu.phenoP1Zoom <- ggplot(data = low.alpha[low.alpha$sim_gen >= 10200,], 
       aes(x = sim_gen, y = meanPhenoP1, group = interaction(mig, mu_base))) + 
    geom_line(aes(color = mig, linetype = mu_base), size = 0.75) + 
    geom_hline(yintercept=1, color = "firebrick", size = 1) + 
    facet_wrap(~ mu_inv, labeller = labeller(mu_inv = inv.labs)) + 
    labs(title = " ",
         y = " ",
         x = "Generation") +
  guides(color = guide_legend(title = "Migration Rate"), 
         linetype = guide_legend(title = "QTN Mutation Rate")) +
  theme_classic() +
  theme(panel.background = element_blank(), 
        strip.background = element_rect(colour = "white", fill = "grey92")) +
  scale_color_manual(values=c("cadetblue1","cadetblue3", "cornflowerblue", "navy")) +
  scale_x_continuous(expand = c(0, 0), limits = c(10000, NA)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(-0.5, 1.2))

grid.arrange(highSigMu.phenoP1, highSigMu.phenoP1Zoom, highSigMu.Fit, highSigMu.LA, nrow = 2, ncol = 2)
grid.arrange(lowSigMu.phenoP1, lowSigMu.phenoP1Zoom, lowSigMu.Fit, lowSigMu.LA, nrow = 2, ncol = 2)

```
## Plot Phenotypes with SD
```{r}
high.alpha$upSD <- high.alpha$meanPhenoP1 + high.alpha$sdPhenoP1
high.alpha$lowSD <-  high.alpha$meanPhenoP1 - high.alpha$sdPhenoP1

highSigMu.lowMig.phenoP1ZoomSD <- ggplot(data = high.alpha[high.alpha$sim_gen >= 10200 & high.alpha$mig == 0.01,],
       aes(x = sim_gen, y = meanPhenoP1, group =  mu_base)) + 
    geom_line(aes(color = mig, linetype = mu_base), size = 0.75) + 
    geom_line(data = high.alpha[high.alpha$sim_gen >= 10200 & high.alpha$mig == 0.01,], 
              aes(x = sim_gen, y = upSD, linetype = mu_base), 
              color = "lightgrey", show.legend = FALSE) +
    geom_line(data = high.alpha[high.alpha$sim_gen >= 10200 & high.alpha$mig == 0.01,], 
              aes(x = sim_gen, y = lowSD, linetype = mu_base,), 
              color = "lightgrey", show.legend = FALSE) +
    geom_hline(yintercept=1, color = "firebrick", size = 0.5) + 
    facet_wrap(~ mu_inv, labeller = labeller(mu_inv = inv.labs)) + 
    labs(title = "",
         y = "Phenotype",
         x = "Generation") +
    guides(color = guide_legend(title = "Migration Rate"), 
           linetype = guide_legend(title = "QTN Mutation Rate")) +
    theme_classic() +
    theme(panel.background = element_blank(), 
          strip.background = element_rect(colour = "white", fill = "grey92")) +
    theme(legend.position = "none") +
    scale_color_manual(values=c("cadetblue3")) +
    scale_x_continuous(expand = c(0, 0), limits = c(10000, NA),
                       breaks = c(10000, 15000, 20000)) + 
    scale_y_continuous(expand = c(0, 0), limits = c(-5, 5))

cols <- c("0.01"="cadetblue3","0.5"="navy", "SD" = "lightgrey")
highSigMuMig.phenoP1ZoomSD <- ggplot(data = high.alpha[high.alpha$sim_gen >= 10200 & high.alpha$mig == 0.5,],
       aes(x = sim_gen, y = meanPhenoP1, group =  mu_base)) + 
    geom_line(aes(color = mig, linetype = mu_base), size = 0.75) + 
    geom_line(data = high.alpha[high.alpha$sim_gen >= 10200 & high.alpha$mig == 0.5,], 
              aes(x = sim_gen, y = upSD, linetype = mu_base ), 
              color = "lightgrey", show.legend = FALSE) +
    geom_line(data = high.alpha[high.alpha$sim_gen >= 10200 & high.alpha$mig == 0.5,], 
              aes(x = sim_gen, y = lowSD, linetype = mu_base ), 
              color = "lightgrey", show.legend = FALSE) +
    geom_hline(yintercept=1, color = "firebrick", size = 0.5) + 
    facet_wrap(~ mu_inv, labeller = labeller(mu_inv = inv.labs)) + 
    labs(title = "",
         y = "",
         x = "Generation") +
    guides(color = guide_legend(title = "Migration Rate"), 
           linetype = guide_legend(title = "QTN Mutation Rate")) +
    theme_classic() +
    theme(panel.background = element_blank(), 
          strip.background = element_rect(colour = "white", fill = "grey92")) +
    scale_color_manual(values=cols, limits = c("0.01", "0.5", "SD"),
                          labels = c("0.01", "0.5",  "SD")) +
    scale_x_continuous(expand = c(0, 0), limits = c(10000, NA), breaks = c(10000, 15000, 20000)) + 
    scale_y_continuous(expand = c(0, 0), limits = c(-5, 5))

grid.arrange(highSigMu.lowMig.phenoP1ZoomSD, highSigMuMig.phenoP1ZoomSD, nrow =1)
 
```
## Inversion length data manipulation
```{r}

df.invTimeFreq <- subset(df.invTime, subset = df.invTime$freq > 0.05)
df.invData$inv_id <- as.numeric(df.invData$inv_id)

df.invAllData <- merge( df.invData, df.invTimeFreq, by.x = c("seed", "inv_id"), by.y = c("seed", "inv_id"), all.y = TRUE)
colnames(df.invAllData)[8] <- "sim_gen"

df.invLength.average <- NULL
inv.sims <- subset(unique.params, subset = unique.params$mu_inv > 0)
for(i in 1:nrow(inv.sims)){
    # create empty variable for putting each data set that should be average
    df.length.average <- NULL
    av.length.seeds <- NULL
    # step through the population dynamics dataframe
    for(j in 1:nrow(df.invAllData)){
      # step through the different seeds that have the unique parameters
      for(k in 1:reps){
        seedCol <- paste("Seed", k, sep="")
        # if that seed is not an NA then do the next step
        if(!is.na(inv.sims[i, seedCol])){
          # if the seed from unique parameters matches the seed of the invTime dataset store it
          if(df.invAllData$seed[j] == inv.sims[i, seedCol]){
           df.length.average <- rbind(df.length.average, df.invAllData[j,])
           av.length.seeds <- unique(c(av.length.seeds, inv.sims[i, seedCol]))
          }
        }
      }
    }
    
   
  ## average columns of interest
   # df.averageSubset <- df.average[,vect.aggCols]
    new.length.average <- aggregate(inv_length~sim_gen, data = df.length.average, FUN = mean)
    new.length.SD <- aggregate(inv_length~sim_gen, data = df.length.average, FUN = sd)
    new.num.inv <- aggregate(inv_length~sim_gen, data = df.length.average, FUN = length)
    new.numQTNs.average <- aggregate(num_qtns~sim_gen, data = df.length.average, FUN = mean)
    df.average.inv <- cbind(new.length.average, new.length.SD[,2],
                            new.numQTNs.average[,2], new.num.inv[,2])
    df.simParams <- data.frame(mu_inv = rep(inv.sims$mu_inv[i], nrow(df.average.inv)), 
                               mig = rep(inv.sims$mig1[i], nrow(df.average.inv)),
                               alpha = rep(inv.sims$alpha[i], nrow(df.average.inv)), 
                               sigmaK = rep(inv.sims$sigmaK[i], nrow(df.average.inv)), 
                               enVar = rep(inv.sims$enVar[i], nrow(df.average.inv)), 
                               mu_base = rep(inv.sims$mu_base[i], nrow(df.average.inv)))
   
    # ## create NA dataframe for columns with no data -- REMOVE CHANGING TO ZEROS NOT NAs
    # df.NAcols <- matrix(NA, nrow = nrow(new.average), ncol = length(vect.noAggCols))
    # colnames(df.NAcols) <- vect.noAggCols
    
    # create seed columns so we can keep track of relevant seed names
    df.Seedcolumns <- NULL
    vect.colNames <- NULL
    #vect.NAseeds <- NULL
    for(m in 1:reps){
      seedCol <- paste("Seed", m, sep = "")
      if(!is.na(av.seeds[m])){
        df.Seedcolumns <- cbind(df.Seedcolumns, rep(av.seeds[m], nrow(df.average.inv)))
      } else {
        df.Seedcolumns <- cbind(df.Seedcolumns, rep(NA, nrow(df.average.inv)))
      }
      vect.colNames <- c(vect.colNames, seedCol)
    }
    colnames(df.Seedcolumns) <- vect.colNames
    
    # finally bind together all the data in the final dataframe
    df.invLength.average <- rbind(df.invLength.average, cbind(df.average.inv, df.simParams, df.Seedcolumns))
    
} # close average for loop
###################################################################
colnames(df.invLength.average)[3] <- "num_qtns"

# convert every column of parameters data to factor 
for(i in 4:ncol(df.invLength.average)){
  df.invLength.average[,i] <- as.factor(as.character(df.invLength.average[,i]))
}
```


### Plot Inversion size and num of qtns
```{r}
## average inversion length 
ggplot(data = df.invLength.average, 
       aes(x = sim_gen, y = inv_length, group = interaction(mig, mu_base))) + 
    geom_line(aes(color = mig, linetype = mu_base), size = 0.75) + 
    facet_wrap(~ alpha) + 
    labs(title = "Average Inversion Length",
           y = "Average Inversion Length",
           x = "Generation") +
  guides(color = guide_legend(title = "Migration Rate"), 
         linetype = guide_legend(title = "QTN Mutation Rate")) +
  theme_classic() +
  theme(panel.background = element_blank(),
        strip.background = element_rect(colour = "white", fill = "grey92")) +
  scale_color_manual(values=c("cadetblue1","cadetblue3", "cornflowerblue", "navy")) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, max(df.invAllData$inv_length)))

## Number of QTNs in inversions
ggplot(data = df.invLength.average, 
       aes(x = sim_gen, y = num_qtns, group = interaction(mig, mu_base))) + 
    geom_line(aes(color = mig, linetype = mu_base), size = 0.75) + 
   facet_wrap(~ alpha) + 
    labs(title = "Average Number of QTNs in Inversions",
           y = "Average Number of QTNs",
           x = "Generation") +
  guides(color = guide_legend(title = "Migration Rate"), 
         linetype = guide_legend(title = "QTN Mutation Rate")) +
  theme_classic() +
  theme(panel.background = element_blank(), 
        strip.background = element_rect(colour = "white", fill = "grey92")) +
  scale_color_manual(values=c("cadetblue1","cadetblue3", "cornflowerblue", "navy")) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, max(df.invLength.average$num_qtns)))

```


```{r}
final.invQTNs <- subset(df.invQTN[df.invQTN$sim_gen == 20000,])
final.invFreq <- subset(df.invAllData[df.invAllData$sim_gen == 20000,])

count <- 1
final.invQTNs$inv_originGen <- NA
for(i in 1:nrow(final.invQTNs)){
  count <- count +1
  for(j in 1:nrow(df.invData)){
    if(final.invQTNs$inv_id[i] == df.invData$inv_id[j]){
      final.invQTNs$inv_originGen[i] <- df.invData$inv_originGen[j]
    }
  }
  print(count)
}

final.invQTNs$BefAft <- NA
for(i in 1:nrow(final.invQTNs)){
  if(final.invQTNs$inv_gen > final.invQTNs$qtn_gen_origin){
    final.invQTNs$BefAft <- "Before"
  } else {
    final.invQTNs$BefAft <- "After"
  }
}
  
beforeAfter <- function(x){
 sum(x == "Before")/length(x) 
}

aggregate(befAft~, data = final.invQTNs, FUN = beforeAfter)
```


## Test Data

```{r}
read.table()



```